### 2016-02-09 HW1/submission/CMW_hw1.tex
Updated the ignore file. Created the tex file for submission.

\documentclass[12pt,a4paper]{article}
\usepackage{amsmath}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{fullpage}
\usepackage[T1]{fontenc}

\author{
  Cheng, Lidens\\
  \texttt{lidenscheng@email.arizona.edu}
  \and
  McClintock, Tom\\
  \texttt{tmcclintock@email.arizona.edu}
  \and
  Wagoner, Erika\\
  \texttt{wagoner47@email.arizona.edu}
}
\title{Astr 513: Homework 1}

\begin{document}
\maketitle

\section{Random Number Generators}

\section{Designing surveys}

\section{Blackbody Distribution}

\end{document}