### 2016-03-29 HW3/HW3.tex
Added file for HW3 and finished problem 1.
Let me know if you see any typos or anything else I should fix/change!

\documentclass[12pt,a4paper]{article}
\usepackage{amsmath}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{fullpage}
\usepackage[T1]{fontenc}
\usepackage{enumerate}
\usepackage[makeroom]{cancel}
\usepackage{hyperref}

\lstdefinestyle{custompython}{
  belowcaptionskip=1\baselineskip,
  breaklines=true,
  frame=L,
  xleftmargin=\parindent,
  language=bash,
  basicstyle=\footnotesize\ttfamily,
  showstringspaces=false,
  %commentstyle=\itshape\color{purple!40!black},
  %keywordstyle=\itshape\color{green!40!black},
  %identifierstyle=\color{blue},
  %stringstyle=\color{orange},
}

\author{
  Cheng, Lidens\\
  \texttt{lidenscheng@email.arizona.edu}
  \and
  McClintock, Tom\\
  \texttt{tmcclintock@email.arizona.edu}
  \and
  Wagoner, Erika\\
  \texttt{wagoner47@email.arizona.edu}
}
\title{Astr 513: Homework 3}

\begin{document}
\maketitle

\section{Linear models of data with correlated uncertainties}
The definition for $\Delta_i$ will not change when $\rho_i \neq 0$, as it does not depend on the uncertainties. So plugging in for $\vec{u}^\top$, $\vec{Z_i}$, and $\cos \theta$, we have
\begin{equation}
 \label{eqn:Delta}
 \Delta_i = \frac{1}{\sqrt{1 + b^2}} \left(\begin{bmatrix}-b & 1\end{bmatrix} \begin{bmatrix}x_i \\ y_i\end{bmatrix} - a\right),
\end{equation}
and simplifying and squaring gives
\begin{equation}
 \label{eqn:Delta2}
 \Delta_i^2 = \frac{\left(y_i - a - b x_i\right)^2}{1 + b^2}.
\end{equation}

However, $\Sigma_i^2$ is defined with the covariance matrix, which will change when the uncertainties are correlated. We have that $\Sigma_i^2 = \vec{u}^\top\, \mathbf{S_i}\, \vec{u}$, and the covariance matrix for correlated uncertainties with $\sigma_{xyi} = \rho_i \sigma_{xi} \sigma_{yi} \neq 0$ is given as, for the data point $(x_i, y_i)$,
\begin{equation}
 \label{eqn:Si}
 \mathbf{S_i} = \begin{bmatrix}\sigma_{xi}^2 & \rho_i \sigma_{xi} \sigma_{yi} \\ \rho_i \sigma_{xi} \sigma_{yi} & \sigma_{yi}^2\end{bmatrix}.
\end{equation}
Thus, we find
\begin{align}
 \Sigma_i^2 &= \frac{1}{1 + b^2} \begin{bmatrix}-b & 1\end{bmatrix} \begin{bmatrix}\sigma_{xi}^2 & \rho_i \sigma_{xi} \sigma_{yi} \\ \rho_i \sigma_{xi} \sigma_{yi} & \sigma_{yi}^2\end{bmatrix} \begin{bmatrix}-b \\ 1\end{bmatrix} \nonumber \\
 &= \frac{1}{1 + b^2} \begin{bmatrix}\sigma_{xi} \left(-b \sigma_{xi} + \rho_i \sigma_{yi}\right) & \sigma_{yi} \left(-b \rho_i \sigma_{xi} + \sigma_{yi}\right)\end{bmatrix} \begin{bmatrix}-b \\ 1\end{bmatrix} \nonumber \\
 &= \frac{1}{1 + b^2} \left(b^2 \sigma_{xi}^2 - b \rho_i \sigma_{xi} \sigma_{yi} - b \rho_i \sigma_{xi} \sigma_{yi} + \sigma_{yi}^2\right) \nonumber \\
 \Rightarrow \Sigma_i^2 &= \frac{\sigma_{yi}^2 -2 b \rho_i \sigma_{xi} \sigma_{yi} + b^2 \sigma_{xi}^2}{1 + b^2}. \label{eqn:Sigma2}
\end{align}

When we then take $\Delta_i^2/2 \Sigma_i^2$, we will see the factor of $1 + b^2$ cancel from both terms, and are left with
\begin{equation}
 \label{eqn:ans1}
 \boxed{\frac{\Delta_i^2}{2 \Sigma_i^2} = \frac{\left(y_i - a - b x_i\right)^2}{2 \left(\sigma_{yi}^2 - 2 b \rho_i \sigma_{xi} \sigma_{yi} + b^2 \sigma_{xi}^2\right)}}
\end{equation}

\section{The Tully-Fisher relation}

\end{document}